Applying Spatio-temporal Scan Statistics and Spatial Autocorrelation Statistics to identify Covid-19 clusters in the world.

### 2021-01-13 Notebooks/paper_covid.rmd
Update paper_covid.rmd

@@

@@

[10] -

>>>>

  When comparing both methods, the Poisson Spatio-Temporal Scan Statistic and the Moran's I, the Scan Statistic was efficient not only to detect emerging *clusters* of Covid-19 cases in space and time by identifying their duration, but also in showing "hidden" *clusters* not visible in LISA maps, such as small european countries like San Marino and Liechtenstein, since the implementation of a Poisson Scan Statistic demands too much computational power, the Moran's I is a better alternative inasmuch as it's easier to calculate and does not demand too much computational power.

@@

 Both the Scan Statistics proposed by Kulldorff, the Global Moran's I and the *LISA* were already used before in studies about spatial and spatio-temporal characteristics of Sars-CoV-2, the Scan Statistic for instance was already used once in studies about the spatio-temporal behavior of Covid-19 in the state of Sergipe in Brazil, while both the Global and Local Moran's I were used in chinese studies regarding the spatial characteristics of the pandemic in China, here these methods showed themselves to be useful while studying the spatial and spatio-temporal characteristics of the virus.

 While the Global Moran's I could be used with great success to identify spatial association in smaller areas affected by the virus and and help in decision making while dealing with local outbreaks, the *LISA* would be useful not only to identify with more precision where the spatial association is occurring but it also should be used when the goal is to control Covid-19 outbreaks in larger areas. The Spatio-Temporal Scan Statistic should be used when trying to identify "alive" *clusters* and when they started, therefore it would be a great tool in decision making when dealing not only with Covid-19 but with other disease outbreaks.

@@
